Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2024-01-12 _targets.Rmd
Complete methods with food-web metrics

```{targets raw_sim}
list(
tar_target(sim_fw_file, here::here("data", "sim_fw"), format = "file"),
  tar_target(sim_fw, {
    x <- readRDS(sim_fw_file)
    x %>%
      mutate(resp_div = 1 - rho)
  }),
  tar_target(sim_fw_df, as.data.frame(sim_fw)),
  tar_target(sim_fw_cons, readRDS(here::here("data", "sim_fw_cons"))),
  tar_target(sim_fw_cons_df, as.data.frame(sim_fw_cons))
  )
```

@@

@@

```{r, message = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(ggcorrplot)
lapply(list.files(here::here("R"), full.names = TRUE), source)
```

@@

@@

```{r}

ggsave_multiple(
  paste0("p_sem_simplified_bivariate_relations", c(".jpg", ".png", ".pdf")),
  plot = p_sem_simplified_bivariate_relations + theme(axis.title = element_text(size = 7)),
  path = here::here("report", "figures"),
  scale = 3.5,
  dpi = 300,
  width = 70,
  height = 70 * .9,
  units = "mm")
```

@@

```{r}
tar_load(fix_fw_stab)
```

```{r}
tar_load(pred_stab_comp_resp_div)
p_pred_sr_resp_div <- pred_stab_comp_resp_div %>%
  filter(response == "stab_com") %>%
  mutate(
    variable = var_replacement()[variable]
    ) %>%
  ggplot(aes(x = resp_div, y = variable_values, fill = slope)) +
  geom_raster() +
  facet_wrap(~variable, scales = "free_y") +
  coord_cartesian(expand = FALSE) +
  scale_fill_gradient2(
    low = resp_div_cat_colors()["Low"],
    mid = "white",
    high = resp_div_cat_colors()["High"],
    midpoint = 0,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill") +
  labs(
    x = var_replacement()["resp_div"],
    y = "Food-web structure",
    fill = "Slope community stability - Richness"
    ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm"),
    plot.margin = margin(l = 10, r = 10, unit = "pt")
  )
ggsave_multiple(
  paste0("p_pred_sr_resp_div", c(".png", ".pdf")),
  plot = p_pred_sr_resp_div,
  path = here::here("report", "figures"),
  scale = 2.2,
  dpi = 400,
  width = 120,
  height = 120 * .45,
  units = "mm")
```

@@

>>>>

```{targets raw_sim}
list(
  tar_target(sim_fw_file, here::here("data", "sim_fw"), format = "file"),
  tar_target(sim_fw, {
    x <- readRDS(sim_fw_file)
    x %>%
      mutate(resp_div = 1 - rho)
  }),
  tar_target(sim_param, readRDS(here::here("data", "sim_param_d2"))),
  tar_target(sim_fw_df, as.data.frame(sim_fw)),
  tar_target(sim_fw_cons, readRDS(here::here("data", "sim_fw_cons"))),
  tar_target(sim_fw_cons_df, as.data.frame(sim_fw_cons))
  )
```

@@

# Table

```{targets}
list(
  tar_target(sim_param_tab,
    sim_param %>%
      mutate(resp_div = 1 - rho) %>%
      select(-A) %>%
      pivot_longer(everything(), names_to = "Parameter", values_to = "values") %>%
      group_by(Parameter) %>%
      summarise(
        u = list(unique(values)),
        n = length(unique(round(values, 2))),
        min = min(values),
        max = max(values)
        )
    ),
  tar_target(tab_sim_param,
    sim_param_tab %>%
      mutate(Values = pmap_chr(list(u = u, min = min, max = max, n = n),
          function(u, min, max, n) {
            if (length(u) < 10) {
              out <- paste0("[", paste(round(u, 2), collapse = ", "), "]")
            } else {
              out <- paste0("(", round(min, 2), ", ", round(max, 2), ")", ", N = ", n)
            }
            out
          }
          )) %>%
      filter(!Parameter %in% c("sim_id", "fw_id", "rho")) %>%
      select(Parameter, Values) %>%
      mutate(Parameter = recode_factor(Parameter, !!!var_replacement()))
  )
)
```

@@

```{r output-load, message = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(ggcorrplot)
library(kableExtra)
lapply(list.files(here::here("R"), full.names = TRUE), source)
```

@@

```{r}
tar_load(sim_param_tab)
```

@@

```{r}
ggsave_multiple(
  paste0("p_sem_simplified_bivariate_relations", c(".jpg", ".png", ".pdf")),
  plot = p_sem_simplified_bivariate_relations + theme(axis.title = element_text(size = 7)),
  path = here::here("report", "figures"),
  scale = 3.5,
  dpi = 300,
  width = 70,
  height = 70 * .9,
  units = "mm")
```

@@

```{r}
tar_load(fix_fw_stab)
```
```{r}
tar_load(pred_stab_comp_resp_div)
```

```{r}
p_pred_sr_resp_div <- pred_stab_comp_resp_div %>%
  filter(response == "stab_com") %>%
  mutate(
    variable = var_replacement()[variable]
    ) %>%
  ggplot(aes(x = resp_div, y = variable_values, fill = slope)) +
  geom_raster() +
  facet_wrap(~variable, scales = "free_y") +
  coord_cartesian(expand = FALSE) +
  scale_fill_gradient2(
    low = resp_div_cat_colors()["Low"],
    mid = "white",
    high = resp_div_cat_colors()["High"],
    midpoint = 0,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill") +
  labs(
    x = var_replacement()["resp_div"],
    y = "Food-web structure",
    fill = "Slope community stability - Richness"
    ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm"),
    plot.margin = margin(l = 10, r = 10, unit = "pt")
  )
ggsave_multiple(
  paste0("p_pred_sr_resp_div", c(".png", ".pdf")),
  plot = p_pred_sr_resp_div,
  path = here::here("report", "figures"),
  scale = 2.2,
  dpi = 400,
  width = 120,
  height = 120 * .45,
  units = "mm")
```

```{r}
p_pred_sr_resp_div_supp <- pred_stab_comp_resp_div %>%
  mutate(
    variable = var_replacement()[variable],
    response = recode_factor(response, !!!var_replacement(), .ordered = TRUE),
    ) %>%
  ggplot(aes(x = resp_div, y = variable_values, fill = slope)) +
  geom_raster() +
  facet_grid(cols = vars(response), rows = vars(variable), scales = "free") +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(
    breaks = c(0, .5, 1),
    labels = c(0, 0.5, 1)
    ) +
  scale_fill_gradient2(
    low = resp_div_cat_colors()["Low"],
    mid = "white",
    high = resp_div_cat_colors()["High"],
    midpoint = 0,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill") +
  labs(
    x = var_replacement()["resp_div"],
    y = "Food-web structure",
    fill = "Slope stability - Richness"
    ) +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(2.0, "cm"),
    plot.margin = margin(l = 10, r = 10, unit = "pt")
  )
ggsave_multiple(
  paste0("p_pred_sr_resp_div_supp", c(".png", ".pdf")),
  plot = p_pred_sr_resp_div_supp,
  path = here::here("report", "figures"),
  scale = 2.4,
  dpi = 400,
  width = 80,
  height = 80 * 1,
  units = "mm")
```

@@

```{r, eval=FALSE}
tar_load(sim_param)
sim_param %<>%
  select(-A)
sim_param %>%
  group_by(h, rho, env_stoch, Z) %>%
  summarise(
    nb_fw = length(unique(fw_id)),
    n = length(fw_id)
  )

sim_param %>%
  group_by(fw_id) %>%
  summarise(n = n())

### 2024-01-12 paper/manuscript.Rmd
Complete methods with food-web metrics

\pagebreak

@@

$$
\epsilon_{i, t} =
  \begin{bmatrix}
     \rho_{11}\sigma^2_e & \rho_{12}\sigma^2_e & \cdots & \rho_{1n}\sigma^2_e \\
     \rho_{21}\sigma^2_e & \rho_{22}\sigma^2_e & \cdots & \rho_{2n}\sigma^2_e \\
     \vdots  & \vdots  & \ddots & \vdots  \\
     \rho_{n1}\sigma^2_e & \rho_{n2}\sigma^2_e & \cdots & \rho_{nn}\sigma^2_e
  \end{bmatrix}
  \begin{bmatrix}
     a_{1,t} \\
     a_{2,t} \\
     \vdots  \\
     a_{n,t}
  \end{bmatrix}
$$

@@

We generated food-web structure using the niche model [@williams_simple_2000]
with an initial species richness from $10$ to $60$, and connectance from $0.02$
to $.32$. We generated food-webs without cannibalistic links and discarded
food-webs that contained disconnected species.

To vary interaction strength and mortality rates in the community,
we varied the Predator-Prey Bodysize Ratio (PPBR) from 1 to 100. We further
varied the functional response of consumer from type II to type III (resp. $h=2$
and $h=3$, eq XX) as it was shown to have a tremendous role on stability
[@brose_allometric_2006]. We set $K = 10$ to ensure that consumers were not
limited by the biomass input from primary producers.

We varied the strength of environmental stochasticity by varying $\sigma_e$ from
0.1 to 0.6, the range of mortality rates observed in
protists [@vasseur_environmental_2007]. We decreased response diversity by
increasing $\rho$ from 0 to 1.

We numerically solved the simulations during 5000 timesteps
and collected the last 500 timesteps. We set $d_t = .1$, and $d_t = .05$ when
instability was detected during the simulations. We set extinction threshold to
$10^{-6}$ and if an extinction happened in the last 500 timesteps, we continued
to run the simulation for another 1000 timesteps until there was no extinction
events. When we found disconnected species at the end of the simulation (i.e. a
primary producer without consumer or a consumer without prey), we removed them
and re-run simulation for another 1000 timesteps.

@@

We measured food-web structure and dynamics that have been shown to be linked
with stability. We collected total biomass, species richness, connectance,
average trophic level weighted by biomass, average omnivory and average per
capita interaction strength.

>>>>

@@

\begin{equation}
\begin{aligned}
\epsilon_{i, t} =
  \begin{bmatrix}
     \rho_{11}\sigma^2_e & \rho_{12}\sigma^2_e & \cdots & \rho_{1n}\sigma^2_e \\
     \rho_{21}\sigma^2_e & \rho_{22}\sigma^2_e & \cdots & \rho_{2n}\sigma^2_e \\
     \vdots  & \vdots  & \ddots & \vdots  \\
     \rho_{n1}\sigma^2_e & \rho_{n2}\sigma^2_e & \cdots & \rho_{nn}\sigma^2_e
  \end{bmatrix}
  \begin{bmatrix}
     a_{1,t} \\
     a_{2,t} \\
     \vdots  \\
     a_{n,t}
  \end{bmatrix}
S_{com} = S_{pop} \times \phi
(\#eq:rho)
\end{aligned}
\end{equation}

@@

```{r}
tar_load(sim_param_tab)
spt <- sim_param_tab
get_sim_par_val <- function(x = "S", data = sim_param_tab) {
  x <- data[data$Parameter == x,][c("min", "max", "n")]
  setNames(unlist(x), c("min", "max", "n"))

}
pS <- get_sim_par_val(x = "S")
pct <- round(get_sim_par_val(x = "ct"), 2)
pZ <- get_sim_par_val(x = "Z")
ph <- get_sim_par_val(x = "h")
pe <- get_sim_par_val(x = "env_stoch")
prd <- get_sim_par_val(x = "resp_div")
pr <- get_sim_par_val(x = "rho")
```
\pagebreak

(ref:eq-rho) (eq. \@ref(eq:rho))

```{r}
tar_read(tab_sim_param) %>%
  kable(
    booktabs = TRUE,
    caption = "Parameter values used in the simulation experiment. Numbers
    between squared brackets display all the values. Two numbers between
    brackets are the minumum and maximal values, N being the number of unique
    values. PPMR: Predator Prey Mass Ratio. We defined response diversity as \\protect$1-\\rho$
    (ref:eq-rho).",
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

We generated food-web structure using the niche model [@williams_simple_2000]
with an initial species richness from $`r pS["min"]`$ to $`r pS["max"]`$, and
connectance from $`r pct["min"]`$ to $`r pct["max"]`$ (Table
\@ref(tab:param)). We generated food-webs without cannibalistic links and
discarded food-webs that contained disconnected species.

We varied the strength of environmental stochasticity by varying $\sigma_e$ from
$`r pe["min"]`$ to $`r pe["max"]`$, the range of mortality rates observed in
protists [@vasseur_environmental_2007]. We decreased response diversity by
increasing $\rho$ from $`r pr["min"]`$ to $`r pr["max"]`$ (i.e. response
diversity being $1 - \rho$) (Fig. \@ref(fig:met)a).

To vary interaction strength and mortality rates (i.e. allometrically
scaled mortality rate $d$) in the community, we varied the Predator-Prey
Bodysize Ratio (PPBR) from $`r pZ["min"]`$ to $`r pZ["max"]`$ (Fig. \@ref(fig:met)b). We further varied
the functional response of consumer from type II to type III (resp.
$`r ph["min"]`$ to $`r ph["max"]`$, eq. \@ref(eq:holling)) as it was shown to
have a tremendous role on stability [@brose_allometric_2006]. We set a global
carrying capacity of $K' = 10$ to ensure that consumers were not limited by the
biomass input from primary producers. We standardised the carrying capacity by
the number of primary producers and by the interspecific competition among
producers to ensure that the effects of species richness on temporal stability
are not driven by the increase in the number of primary producers, thereby
trivially increase biomass input for consumers. We then standardised $K'$ such as
$K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K'$ [@ives_stability_2000].

We numerically solved the stochastic differential equation systems during 5000
timesteps and collected the last 500 timesteps. We set $d_t = .1$, and $d_t =
.05$ when instability was detected during the simulations due to stiff changes
in biomass. We set extinction threshold to $10^{-6}$ and if an extinction
happened in the last 500 timesteps, we continued to run the simulation for
another 1000 timesteps until there was no extinction events. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we removed them and re-run the
simulation for another 1000 timesteps.

@@

We measured food-web structure and dynamics that have been shown to be linked
with stability. We collected total biomass, species richness, connectance,
average trophic level weighted by biomass, average omnivory and average per
capita interaction strength. Connectance was computed as $C = L / S(S-1)$, $L$
being the number of trophic links and $S$ being the number of species in the
community. Species trophic level was computed recursevily from the bottom of the
food-web, such as the trophic level of a consumer was equal to the average
trophic level of its resources added to 1, the trophic level of primary
producers being set to 1 [@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of square of its
resource trophic levels weighted by the relative preference of the consumer for
this resource: $\sum_i (\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$ [@christensen_ecopath_1992].
Interaction strength was computed was quantified as the biomass fluxes going
from a resource to a consumer, such as: $I = x_i y_i B_k F_{ij}$ (eq.
\@ref(eq:cons)), and averaged over time. Per capita interaction strength was
obtained by dividing average interaction strength by the average biomass of the
consumer. Finally, the average interaction strength of a community was computed
as the average per capita interaction strength at the exclusion of null
interaction strengths (i.e. non present trophic interactions).
